### 2016-05-17 index.Rmd
Use a single busy indicator

---
title: "Volatility and the Buyback Anomaly: Interactive Tool"
author: "[Theodoros Evgeniou](http://faculty.insead.edu/theodoros-evgeniou/), [Enric Junqué de Fortuny](http://ciri.be/), [Nick Nassuphis](https://www.linkedin.com/in/nick-nassuphis-8029372), and [Theo Vermaelen](http://faculty.insead.edu/theo-vermaelen/)"
output:  
 html_document:
 fig_width: 6
 fig_height: 2.5
 fig_align: center
 in_header: header.html
 runtime: shiny
runtime: shiny
---

```{r, eval=TRUE, echo = FALSE,message=FALSE}
# Copyright 2015, INSEAD
# by T. Evgeniou, Enric Junque de Fortuny, Nick Nassuphis, Theo Vermaelen 
# Dual licensed under the MIT or GPL Version 2 licenses.  
```

```{r, eval=TRUE, echo = FALSE,message=FALSE}
# need this to avoid loading the R libraries when the tool is deployed on shinyapps (needs to be 0 when run localy, 1 when deployed on a shinyapps server)

run_shiny_tool = 1

```

```{r, eval=TRUE, echo = FALSE,message=FALSE}
# rm(list=ls()) # Clean up the memory, if we want to rerun from scratch
load("bb_issuersTOOL.Rdata")

if(run_shiny_tool) suppressPackageStartupMessages({
  library(htmlwidgets)
  library(shiny)
  library(shinyjs)
  library(shinysky) #  devtools::install_github("AnalytixWare/ShinySky")
  library(dygraphs)
  library(Rcpp)
  library(RcppArmadillo)
  library(stringr)
  library(gtools)
  library(timeDate)
})

div(
  style = "width:32px; height:32px; position:fixed; top:0; bottom:0; left:0; right:0; margin: auto; z-index:9999",
  busyIndicator("", wait=111)
)

source("lib_helpers.R", local = TRUE)
source("latex_code.R")
source("ff_industries_sic.R")
source("Paper_global_parameters.R")
source("heatmapOutput.R") 
source("ff_industries_sic.R")

report_months_cal = c(as.numeric(reported_times[1:(length(reported_times)-1)]),1)
BUYBACK_DATA_TOOL$None <- 1:length(BUYBACK_DATA_TOOL$Prior.Returns.Score)

value.weights = rep(1,length(BUYBACK_DATA_TOOL$Prior.Returns.Score))
acceptaple_features = setdiff(names(BUYBACK_DATA_TOOL),c("Event.Date", "Industry_name","Event.Size"))
```

<br>

**NOTE:** Please always wait until all tables load (and "fade in")...

<hr>

This is an interactive tool for the article [Volatility and the Buyback Anomaly](http://tevgeniou.github.io/BuybacksIssuers/). Users can explore the effects of various parameters as well as data filtering choices. For any given choices, users can generate and download a new **customized pdf version of the paper** (work in progress) by clicking the button below at any time.

**Note that it takes a few minutes (possibly more than five) to compile the new paper. Please wait until the option to download the paper, becomes available.**

```{r, fig.width=6, fig.height=5,echo=FALSE}
useShinyjs(rmd=TRUE)
fluidPage(
  shiny::actionButton("compileReport","Compile Paper",icon=icon("spinner")),
  shiny::downloadButton("downloadReport", "Download Paper")
)
report.data <- paste0("session.", session$token, ".report.data.Rdata")
report.script <- paste0("session.", session$token, ".report.script.R")
report.source <- paste0("session.", session$token, ".report.Rnw")
report.file <- paste0("session.", session$token, ".report.pdf")
observeEvent(input$compileReport, {
  disable("compileReport")
  with(report.list <- new.env(), {
    startdate <- isolate(input$startdate)
    enddate <- isolate(input$enddate)
    market_cap_min <- isolate(input$market_cap_min)
    market_cap_max <- isolate(input$market_cap_max)
    characteristic <- isolate(input$characteristic)
    characteristic_single_analysis <- isolate(input$characteristic_single_analysis)
    quantile_used <- isolate(input$quantile_used)
    characteristic_single_analysis_sign <- isolate(input$characteristic_single_analysis_sign)
    high_EU_index_thres <- isolate(input$high_EU_index_thres)
    low_EU_index_thres <- isolate(input$low_EU_index_thres)
    quantile_used_ind <- isolate(input$quantile_used_ind)
    quantile_used_all <- isolate(input$quantile_used_all)
    characteristic_irats1 <- isolate(input$characteristic_irats1)
    characteristic_irats2 <- isolate(input$characteristic_irats2)
    characteristic_irats3 <- isolate(input$characteristic_irats3)
    characteristic_irats4 <- isolate(input$characteristic_irats4)
    characteristic_irats5 <- isolate(input$characteristic_irats5)
    characteristic_irats6 <- isolate(input$characteristic_irats6)
    characteristic_irats1_sign <- isolate(input$characteristic_irats1_sign)
    characteristic_irats2_sign <- isolate(input$characteristic_irats2_sign)
    characteristic_irats3_sign <- isolate(input$characteristic_irats3_sign)
    characteristic_irats4_sign <- isolate(input$characteristic_irats4_sign)
    characteristic_irats5_sign <- isolate(input$characteristic_irats5_sign)
    characteristic_irats6_sign <- isolate(input$characteristic_irats6_sign)
    industry <- isolate(input$industry)
  })
  for (x in ls(report.list)) if (is.null(get(x, report.list))) warning(paste0("parameter '", x, "' is NULL"), call.=FALSE, immediate.=TRUE)
  save(report.list, file=report.data)
  file.copy("toolreport.Rnw", report.source)
  cat(paste0("load ('", report.data, "')\n"), file=report.script)
  cat(paste0("knitr::knit2pdf('", report.source,"', quiet=TRUE)\n"), file=report.script, append=TRUE)
  system(paste0("Rscript --vanilla ", report.script), wait=FALSE)
})
observe({
  invalidateLater(1000, session)
  toggleState("downloadReport", condition=file.exists(report.file))
})
output$downloadReport <- downloadHandler(
  filename = "report.pdf",
  content = function(file) {
    file.copy(report.file, file)
    file.remove(list.files(pattern=paste0("session.", session$token)))
    enable("compileReport")
    disable("downloadReport")
  }
)
```

<hr>

### Dataset Filtering

You can select some filters for the dataset. All analyses below will be done only for the events that are selected.

Every time you change these filters you may **need to wait a few seconds** until all results of this tool are updated (all tables/figures "fade in").

<br>

Select the time period for which you would like to do the analyses below:

<br>

```{r, eval=TRUE, echo = FALSE}
dateInput("startdate", "Starting Date:", 
          value = min(BUYBACK_DATA_TOOL$Event.Date) - 10, min = min(BUYBACK_DATA_TOOL$Event.Date), max = max(BUYBACK_DATA_TOOL$Event.Date)-3*370,
          startview = "year",
          width='100%')
```
<br>
```{r, eval=TRUE, echo = FALSE}
dateInput("enddate", "End Date (allow at least 3 years from the starting day you selected - else it will automatically use data up to 3 years after your start date):", 
          value = max(BUYBACK_DATA_TOOL$Event.Date) + 10, 
          min = min(BUYBACK_DATA_TOOL$Event.Date)+3*370, 
          max = max(BUYBACK_DATA_TOOL$Event.Date)+10,startview = "year",
          width='100%')
```

<br>

Selet the minimum size of the firms (in $millions) for which you would like to do the analyses below. The maximum selected should be larger than the minimum, else the maximum is automatically set to infinity:

```{r, eval=TRUE, echo = FALSE}
selectInput("market_cap_min","Enter the Minimum Market Cap allowed (in millions):", choices = as.character(c(0,50*(1:20), 500*(1:20))), selected = "0", width = "1000px")
```
<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("market_cap_max","Enter the Maximum Market Cap allowed (in millions):", choices = as.character(c(0,50*(1:20), 500*(1:20), 10e10)), selected = as.character(10e10), width = "1000px")
```

We can also see the effects of removing some industries: we can select industries to remove here (make sure there are enough data left for the analysis):

<br>

```{r, eval=TRUE, echo = FALSE}
selectInput("industry", "Select Industries to remove:", choices = unique(BUYBACK_DATA_TOOL$Industry_name),selected = NULL,multiple=TRUE,width="100%") 
```
<br>

```{r, eval=TRUE, echo = FALSE}
DATASET_NEW <- reactive({
  starting = input$startdate 
  ending = input$enddate   
  if (ending - starting < 3*365)
    ending = starting + 3*365

  market_cap_min = as.numeric(input$market_cap_min)
  market_cap_max = as.numeric(input$market_cap_max)
  if (market_cap_max < market_cap_min)
    market_cap_max = ceiling(max(BUYBACK_DATA_TOOL$Market.Cap))

  useonly_industry = !(BUYBACK_DATA_TOOL$Industry_name %in% input$industry)

  useonly_report = which(BUYBACK_DATA_TOOL$Event.Date >= starting & BUYBACK_DATA_TOOL$Event.Date <= ending &
                           scrub(BUYBACK_DATA_TOOL$Market.Cap) >= market_cap_min & 
                           scrub(BUYBACK_DATA_TOOL$Market.Cap) <= market_cap_max &
                           useonly_industry)

  list(
    BUYBACK_DATA_TOOL=BUYBACK_DATA_TOOL[useonly_report,],
    returns_by_event_monthly = returns_by_event_monthly[,useonly_report],
    DatesMonth  = DatesMonth[,useonly_report],
    value.weights = value.weights[useonly_report],
    subset = useonly_report
  )
})

```

<hr>

### Descriptive Statistics

This tool allows the user to explore descriptive statistics of some of the key firm characteristics. When the name of the firm characteristics ends with ".Score" it is the percentile (0 to 1) relative to all firms in the CRSP universe at the time of the buyback announcement, as discussed in the paper. Otherwise it is the raw value of the firm characteristic.

<br>

```{r, eval=TRUE, echo = FALSE}
fluidPage(
  selectInput("characteristic", "Select Firm Characteristic:", choices = acceptaple_features,selected = "EU.Index"),
  mainPanel(
    navlistPanel(
      tabPanel("Histogram",
               renderPlot({
                 feature = DATASET_NEW()$BUYBACK_DATA_TOOL[,which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic)]
                 feature = feature[!is.na(feature)]
                 if (!is.numeric(feature))
                   feature = 1
                 hist(feature, density = 100, breaks = 100)
               })),
      tabPanel("Summary Statistics",
               renderDataTable({
                 feature = DATASET_NEW()$BUYBACK_DATA_TOOL[,which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic)]
                 feature = feature[!is.na(feature)]
                 if (length(unique(feature)) > 20 & is.numeric(feature)){
                   res = t(round(as.matrix(summary(feature)),2))
                   colnames(res) <- c("Minimum", "1st Quantile", "Median", "Mean", "3rd Quantile", "Max")
                 }
                 if (is.character(feature) | length(unique(feature)) <= 20){
                   tmp = table(feature)
                   res = t(as.matrix(tmp))
                 }
                 as.data.frame(res)
                 #m1<-gvisTable(show_data,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
                 #print(m1,'chart')
               })
      ),
      widths = c(3, 9)),
    style='width: 100%;'))
```

<hr>

### Buybacks and Firm Characteristics

The article relates 5 firm characteristics (pre-buyback-announcement returns, idiosyncratic volatility, and volatility, as well as BE/ME and Firm Size). This tool allows the user to study the effects of combining any of these firm characteristics, as well as others not discussed in the paper. We consider here high/low events defined using the quantiles of the selected event charactetistic.

```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_single_analysis", "Select Firm Characteristic to Define High/Low Firms:", choices = acceptaple_features,selected = "EU.Index", width = "1000px")

selectInput("quantile_used","Enter the quantile to use to define High (>=) and Low (<=) Firms:", choices = 0.02*(1:20), selected = "0.1", width = "1000px")

checkboxInput("characteristic_single_analysis_sign", "If you would like to change the sign of the firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")

feature_IRATStable <- reactive({
  thesign = ifelse(input$characteristic_single_analysis_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]
  thefeatureQ = thesign*BUYBACK_DATA_TOOL[[which(names(BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]  
  quantile_used_ind = as.numeric(input$quantile_used)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= quantile(scrub(thefeatureQ[!is.na(thefeatureQ)]),1-quantile_used_ind)
  Low_feature_events  = !is.na(thefeature) & scrub(thefeature) <= quantile(scrub(thefeatureQ[!is.na(thefeatureQ)]),quantile_used_ind)

  res = round(cbind(
    car_table(DATASET_NEW()$returns_by_event_monthly[, Low_feature_events],
              DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[ Low_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model)$results,
    car_table(DATASET_NEW()$returns_by_event_monthly[, High_feature_events],
              DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[High_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model)$results
  ),2)[reported_times,]

  colnames(res) <- c("Low: CAR", "t-stat","p-value", "High: CAR", "t-stat","p-value")
  as.data.frame(res)
})

feature_CALtable <- reactive({
  thesign = ifelse(input$characteristic_single_analysis_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]
  thefeatureQ = thesign*BUYBACK_DATA_TOOL[[which(names(BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]  
  quantile_used_ind = as.numeric(input$quantile_used)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= quantile(scrub(thefeatureQ[!is.na(thefeatureQ)]),1-quantile_used_ind)
  Low_feature_events  = !is.na(thefeature) & scrub(thefeature) <= quantile(scrub(thefeatureQ[!is.na(thefeatureQ)]),quantile_used_ind)

  res = round(cbind(
    calendar_table(DATASET_NEW()$returns_by_event_monthly[, Low_feature_events],
                   DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[Low_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model, value.weights = DATASET_NEW()$value.weights[Low_feature_events])$results,
    calendar_table(DATASET_NEW()$returns_by_event_monthly[,High_feature_events],
                   DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[High_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model, value.weights = DATASET_NEW()$value.weights[High_feature_events])$results
  ),2)[reported_times,]

  colnames(res) <- c("Low: CAL", "t-stat","p-value", "High: CAL", "t-stat","p-value")
  as.data.frame(res)
})

High_feature_Hedged_react <- reactive({
  thesign = ifelse(input$characteristic_single_analysis_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]
  thefeatureQ = thesign*BUYBACK_DATA_TOOL[[which(names(BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]  
  quantile_used_ind = as.numeric(input$quantile_used)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= quantile(scrub(thefeatureQ[!is.na(thefeatureQ)]),1-quantile_used_ind)

  High_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", High_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(High_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)
})

Low_feature_Hedged_react <- reactive({
  thesign = ifelse(input$characteristic_single_analysis_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]
  thefeatureQ = thesign*BUYBACK_DATA_TOOL[[which(names(BUYBACK_DATA_TOOL) == input$characteristic_single_analysis)]]  
  quantile_used_ind = as.numeric(input$quantile_used)

  Low_feature_events  = !is.na(thefeature) & scrub(thefeature) <= quantile(scrub(thefeatureQ[!is.na(thefeatureQ)]),quantile_used_ind)

  Low_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", Low_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(Low_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)  
})

```

<br>

```{r, eval=TRUE, echo = FALSE}
fluidPage(
  mainPanel(
    navlistPanel(
      tabPanel("Five-Factor IRATS Cumulative Abnormal Returns",
               renderTable({
                 feature_IRATStable()
               })),
      tabPanel("Five-Factor Calendar Method Monthly Abnormal Returns",
               renderTable({
                 feature_CALtable()
               })),
      tabPanel("Cumulative Abnormal Returns,High, 12m hold",
               renderDygraph({
                 pnl_plot_interactive(High_feature_Hedged_react())
               })),
      tabPanel("Cumulative Abnormal Returns,Low, 12m hold",
               renderDygraph({
                 pnl_plot_interactive(Low_feature_Hedged_react())
               })),
      tabPanel("Monthly and Yearly Returns, High, 12m hold",
               renderUI(
                 HTML(renderHeatmapX(pnl_matrix(High_feature_Hedged_react())))
               )),
      tabPanel("Monthly and Yearly Returns, Low, 12m hold",
               renderUI(
                 HTML(renderHeatmapX(pnl_matrix(Low_feature_Hedged_react())))
               )),
      widths = c(3, 9)),style='width: 100%;'))
```

<hr>

### The EU-Index

For the subset of events selected the High/Low EU index events (for the selected EU thresholds below - you may need to change them depending on the sample of events used) have these abnormal returns:

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("high_EU_index_thres","Enter the EU index threshold above which (>=) we define the High EU-Index firms (need minimum 10 firms High EU firms):", choices = sort(unique(BUYBACK_DATA_TOOL$EU.Index)), selected = "5", width = "1000px")
```

```{r, eval=TRUE, echo = FALSE}
selectInput("low_EU_index_thres","Enter the EU index threshold below which (<=) we define the Low EU-Index firms (need minimum 10 firms Low EU firms):", choices = sort(unique(BUYBACK_DATA_TOOL$EU.Index)), selected = "1", width = "1000px")
```

<br>

```{r, eval=TRUE, echo = FALSE}
EU_index_table <- reactive({
  thefeature = DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == "EU.Index")]]
  res = matrix(table(thefeature),nrow=1)
  colnames(res) <- paste("EU Index", names(table(thefeature)), sep=": ")
  rownames(res) <- "# of Events"
  res
})

feature_IRATStable_industry <- reactive({
  thefeature = DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == "EU.Index")]]

  high_EU_index_thres = as.numeric(input$high_EU_index_thres)
  low_EU_index_thres = as.numeric(input$low_EU_index_thres)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= high_EU_index_thres
  if (sum(High_feature_events) < 10)
    High_feature_events = !is.na(thefeature) & scrub(thefeature) >= high_EU_index_thres-1

  Low_feature_events  = !is.na(thefeature) & scrub(thefeature) <= low_EU_index_thres
  if (sum(Low_feature_events) < 10)
    Low_feature_events = !is.na(thefeature) & scrub(thefeature) <= Low_feature_events+1

  res = round(cbind(
    car_table(DATASET_NEW()$returns_by_event_monthly[,Low_feature_events],
              DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[Low_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model)$results,
    car_table(DATASET_NEW()$returns_by_event_monthly[,High_feature_events],
              DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[High_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model)$results
  ),2)[reported_times,]
  colnames(res) <- c("Low EU: CAR", "t-stat","p-value", "High EU: CAR", "t-stat","p-value")
  as.data.frame(res)
})

feature_CALtable_industry <- reactive({
  thefeature = DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == "EU.Index")]]

  high_EU_index_thres = as.numeric(input$high_EU_index_thres)
  low_EU_index_thres = as.numeric(input$low_EU_index_thres)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= high_EU_index_thres
  if (sum(High_feature_events) < 10)
    High_feature_events = !is.na(thefeature) & scrub(thefeature) >= high_EU_index_thres-1

  Low_feature_events  = !is.na(thefeature) & scrub(thefeature) <= low_EU_index_thres
  if (sum(Low_feature_events) < 10)
    Low_feature_events = !is.na(thefeature) & scrub(thefeature) <= Low_feature_events+1

  res = round(cbind(
    calendar_table(DATASET_NEW()$returns_by_event_monthly[,Low_feature_events],
                   DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[Low_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model, value.weights = DATASET_NEW()$value.weights[Low_feature_events])$results,
    calendar_table(DATASET_NEW()$returns_by_event_monthly[,High_feature_events],
                   DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[High_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model, value.weights = DATASET_NEW()$value.weights[High_feature_events])$results
  ),2)[reported_times,]
  colnames(res) <- c("Low EU: CAL", "t-stat","p-value", "High EU: CAL", "t-stat","p-value")
  as.data.frame(res)
})

fluidPage(
  mainPanel(
    navlistPanel(
      tabPanel("Number of Firms per EU Index Value",
               renderTable({
                 EU_index_table()
               })),
      tabPanel("Five-Factor IRATS Cumulative Abnormal Returns",
               renderTable({
                 feature_IRATStable_industry()
               })),
      tabPanel("Five-Factor Calendar Method Monthly Abnormal Returns",
               renderTable({
                 feature_CALtable_industry()
               })),
      widths = c(3, 9)),style='width: 100%;'))
```

<hr>

### Customized Buyback Indices

Much like the EU-index in the paper, one can create other indices by combining different firm characteristics as needed. You can select your own firm characteristics below and also indicate whether you would like to reverse the sign of this firm characteristic (Low will be High and vice versa).

<br>
```{r, eval=TRUE, echo = FALSE}
fluidPage( 
  h4("Firm Characteristics of the new index"),
  fluidRow(
    column(6,
           selectInput("characteristic_irats1", "First Firm Characteristic:", choices = acceptaple_features,selected = "EU.Index", width = "425px"),
           checkboxInput("characteristic_irats1_sign", "Reverse", value = FALSE),
           selectInput("characteristic_irats2", "Second Firm Characteristic:", choices = acceptaple_features,selected ="None", width = "425px"),
           checkboxInput("characteristic_irats2_sign", "Reverse", value = FALSE),
           selectInput("characteristic_irats3", "Third Firm Characteristic:", choices = acceptaple_features,selected = "None", width = "425px"),
           checkboxInput("characteristic_irats3_sign", "Reverse", value = FALSE)
    ),
    column(6,
           selectInput("characteristic_irats4", "Fourth Firm Characteristic:", choices = acceptaple_features,selected = "None", width = "425px"),
           checkboxInput("characteristic_irats4_sign", "Reverse", value = FALSE),
           selectInput("characteristic_irats5", "Fifth Firm Characteristic:", choices = acceptaple_features,selected = "None", width = "425px"),
           checkboxInput("characteristic_irats5_sign", "Reverse", value = FALSE),
           selectInput("characteristic_irats6", "Sixth Firm Characteristic:", choices = acceptaple_features,selected = "None", width = "425px"),
           checkboxInput("characteristic_irats6_sign", "Reverse", value = FALSE)
    )
  )
)
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("quantile_used_ind","Enter the quantile to use to define the 0 (Low, <=), 1 (middle), 2 (High, >=) scores for each of the selected characteristics:", choices = 0.02*(1:20), selected = "0.1", width = "1000px")
```

<br>
```{r, eval=TRUE, echo = FALSE}

selectInput("quantile_used_all","Enter the quantile to use to define High (>=) and Low (<=) Firms for the total index created (sum of the scores of the firm characteristics selected):", choices = 0.02*(1:20), selected = "0.1", width = "1000px")
```

```{r, eval=TRUE, echo = FALSE}
selected_features = reactive({
  paste(setdiff(unique(c(input$characteristic_irats1,input$characteristic_irats2,input$characteristic_irats3,input$characteristic_irats4,input$characteristic_irats5,input$characteristic_irats6)), "None"), sep=",")
})

```

<br>

<br> 
Here are the results for your new customized index:

<br>
```{r, eval=TRUE, echo = FALSE}

Index_Score_rect <- reactive({

  thesign = ifelse(input$characteristic_irats1_sign, -1, +1)
  thefeature1 = (thesign==-1) + thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_irats1)]]  
  thefeature1Q = (thesign==-1) + thesign*BUYBACK_DATA_TOOL[[which(names(BUYBACK_DATA_TOOL) == input$characteristic_irats1)]]  
  High_feature_events1 = (scrub(thefeature1) >= quantile(thefeature1Q[!is.na(thefeature1Q)],1-as.numeric(input$quantile_used_ind)) & !is.na(thefeature1))
  Low_feature_events1 = (scrub(thefeature1) <= quantile(thefeature1Q[!is.na(thefeature1Q)],as.numeric(input$quantile_used_ind)) & !is.na(thefeature1))

  thefeature2 <- thefeature3 <- thefeature4 <- thefeature5 <- thefeature6 <- rep(0,length(High_feature_events1))
  High_feature_events2 <- High_feature_events3 <- High_feature_events4 <- High_feature_events5 <- High_feature_events6 <- rep(FALSE,length(High_feature_events1))
  Low_feature_events2 <- Low_feature_events3 <- Low_feature_events4 <- Low_feature_events5 <- Low_feature_events6 <- 
    rep(FALSE,length(Low_feature_events1))

  if (input$characteristic_irats2 != "None"){
    thesign = ifelse(input$characteristic_irats2_sign, -1, +1)
    thefeature2 = (thesign==-1) + thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_irats2)]]  
    thefeature2Q = (thesign==-1) + thesign*BUYBACK_DATA_TOOL[[which(names(BUYBACK_DATA_TOOL) == input$characteristic_irats2)]]  
    High_feature_events2 = (scrub(thefeature2) >= quantile(thefeature2Q[!is.na(thefeature2Q)],1-as.numeric(input$quantile_used_ind)) & !is.na(thefeature2))
    Low_feature_events2 = (scrub(thefeature2) <= quantile(thefeature2Q[!is.na(thefeature2Q)],as.numeric(input$quantile_used_ind)) & !is.na(thefeature2))
  }
  if (input$characteristic_irats3 != "None"){
    thesign = ifelse(input$characteristic_irats3_sign, -1, +1)
    thefeature3 = (thesign==-1) + thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_irats3)]]  
    High_feature_events3 = (scrub(thefeature3) >= quantile(thefeature3[!is.na(thefeature3)],1-as.numeric(input$quantile_used_ind)) & !is.na(thefeature3))
    Low_feature_events3 = (scrub(thefeature3) <= quantile(thefeature3[!is.na(thefeature3)],as.numeric(input$quantile_used_ind)) & !is.na(thefeature3))
  }
  if (input$characteristic_irats4 != "None"){
    thesign = ifelse(input$characteristic_irats4_sign, -1, +1)
    thefeature4 = (thesign==-1) + thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_irats4)]]  
    High_feature_events4 = (scrub(thefeature4) >= quantile(thefeature4[!is.na(thefeature4)],1-as.numeric(input$quantile_used_ind)) & !is.na(thefeature4))
    Low_feature_events4 = (scrub(thefeature4) <= quantile(thefeature4[!is.na(thefeature4)],as.numeric(input$quantile_used_ind)) & !is.na(thefeature4))
  }
  if (input$characteristic_irats5 != "None"){
    thesign = ifelse(input$characteristic_irats5_sign, -1, +1)
    thefeature5 = (thesign==-1) + thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_irats5)]]  
    High_feature_events5 = (scrub(thefeature5) >= quantile(thefeature5[!is.na(thefeature5)],1-as.numeric(input$quantile_used_ind)) & !is.na(thefeature5))
    Low_feature_events5 = (scrub(thefeature5) <= quantile(thefeature5[!is.na(thefeature5)],as.numeric(input$quantile_used_ind)) & !is.na(thefeature5))
  }
  if (input$characteristic_irats6 != "None"){
    thesign = ifelse(input$characteristic_irats6_sign, -1, +1)
    thefeature6 = (thesign==-1) + thesign*DATASET_NEW()$BUYBACK_DATA_TOOL[[which(names(DATASET_NEW()$BUYBACK_DATA_TOOL) == input$characteristic_irats6)]]  
    High_feature_events6 = (scrub(thefeature5) >= quantile(thefeature5[!is.na(thefeature5)],1-as.numeric(input$quantile_used_ind)) & !is.na(thefeature5))
    Low_feature_events6 = (scrub(thefeature5) <= quantile(thefeature5[!is.na(thefeature5)],as.numeric(input$quantile_used_ind)) & !is.na(thefeature5))
  }

  Index_score = 
    2*High_feature_events1 + 1*(!High_feature_events1 & !Low_feature_events1) + 
    (input$characteristic_irats2 != "None")*(2*High_feature_events2 + 1*(!High_feature_events2 & !Low_feature_events2) ) + 
    (input$characteristic_irats3 != "None")*(2*High_feature_events3 + 1*(!High_feature_events3 & !Low_feature_events3) ) + 
    (input$characteristic_irats4 != "None")*(2*High_feature_events4 + 1*(!High_feature_events4 & !Low_feature_events4) ) + 
    (input$characteristic_irats5 != "None")*(2*High_feature_events5 + 1*(!High_feature_events5 & !Low_feature_events5) ) + 
    (input$characteristic_irats6 != "None")*(2*High_feature_events6 + 1*(!High_feature_events6 & !Low_feature_events6) )

  Index_score

})

feature_IRATStable_index <- reactive({
  thefeature = Index_Score_rect()
  High_feature_events = (scrub(thefeature) >= quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used_all)) & !is.na(thefeature))
  Low_feature_events = (scrub(thefeature) <= quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used_all)) & !is.na(thefeature))

  res = round(cbind(
    car_table(DATASET_NEW()$returns_by_event_monthly[, Low_feature_events],
              DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[ Low_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model)$results,
    car_table(DATASET_NEW()$returns_by_event_monthly[, High_feature_events],
              DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[High_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model)$results
  ),2)[reported_times,]
  colnames(res) <- c("Low: CAR", "t-stat","p-value", "High: CAR", "t-stat","p-value")
  res
})

feature_CALtable_index <- reactive({
  thefeature = Index_Score_rect()
  High_feature_events = (scrub(thefeature) >= quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used_all)) & !is.na(thefeature))
  Low_feature_events = (scrub(thefeature) <= quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used_all)) & !is.na(thefeature))

  res = round(cbind(
    calendar_table(DATASET_NEW()$returns_by_event_monthly[, Low_feature_events],
                   DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[Low_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model, value.weights = DATASET_NEW()$value.weights[Low_feature_events])$results,
    calendar_table(DATASET_NEW()$returns_by_event_monthly[,High_feature_events],
                   DATASET_NEW()$BUYBACK_DATA_TOOL$Event.Date[High_feature_events],Risk_Factors_Monthly,formula_used = five_factor_model, value.weights = DATASET_NEW()$value.weights[High_feature_events])$results
  ),2)[reported_times,]

  colnames(res) <- c("Low: CAL", "t-stat","p-value", "High: CAL", "t-stat","p-value")
  res
})

High_feature_Hedged_react_index <- reactive({
  thefeature = Index_Score_rect()
  High_feature_events = which(scrub(thefeature) >= quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used_all)) & !is.na(thefeature))
  High_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", High_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(High_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)
})

Low_feature_Hedged_react_index <- reactive({
  thefeature = Index_Score_rect()
  Low_feature_events = which(scrub(thefeature) <= quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used_all)) & !is.na(thefeature))
  Low_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", Low_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(Low_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)  
})

```

```{r, eval=TRUE, echo = FALSE}
fluidPage(
  mainPanel(
    navlistPanel(
      tabPanel("Histogram of the New Index",
               renderPlot({
                 feature = Index_Score_rect()
                 feature = feature[!is.na(feature)]
                 if (!is.numeric(feature))
                   feature = 1
                 hist(feature, density = 100, breaks = 20,main="Index Histogram",xlab="Score")
               })),
      tabPanel("Five-Factor IRATS Cumulative Abnormal Returns",
               renderTable({
                 feature_IRATStable_index()
               })),
      tabPanel("Five-Factor Calendar Method Monthly Abnormal Returns",
               renderTable({
                 feature_CALtable_index()
               })),
      tabPanel("Cumulative Abnormal Returns,High, 12m hold",
               renderDygraph({
                 pnl_plot_interactive(High_feature_Hedged_react_index())
               })),
      tabPanel("Cumulative Abnormal Returns,Low, 12m hold",
               renderDygraph({
                 pnl_plot_interactive(Low_feature_Hedged_react_index())
               })),
      tabPanel("Monthly and Yearly Returns, High, 12m hold",
               renderUI(
                 HTML(renderHeatmapX(pnl_matrix(High_feature_Hedged_react_index())))
               )),
      tabPanel("Monthly and Yearly Returns, Low, 12m hold",
               renderUI(
                 HTML(renderHeatmapX(pnl_matrix(Low_feature_Hedged_react_index())))
               )),
      widths = c(3, 9)),style='width: 100%;'))
```

<br>
<hr>
As noted above **you can interactively change various report parameters throughout this tool, and for any of your choices create a customized paper using the button at the begining of this page at any time.**
<br>

<hr>

<br>

<br>
<br>